Hillary Clinton Emails

### 2016-04-12 presentation/presentation.tex
Added data and MDS/clustering to presentation.

% \documentclass[14pt]{beamer}
\documentclass[handout,14pt]{beamer}

@@

\usetheme{Madrid}
% \usecolortheme{}

@@

\begin{frame}{Latent Dirichlet Allocation}

\begin{figure}
\centering
\begin{tikzpicture}
\tikzstyle{main}=[circle, minimum size = 10mm, thick, draw =black!80, node distance = 16mm]
\tikzstyle{connect}=[-latex, thick]
\tikzstyle{box}=[rectangle, draw=black!100]
  \node[main, fill = white!100] (alpha) [label=below:$\alpha$] { };
  \node[main] (theta) [right=of alpha,label=below:$\theta$] { };
  \node[main] (z) [right=of theta,label=below:z] {};
  \node[main] (beta) [above=of z,label=below:$\beta$] { };
  \node[main, fill = black!10] (w) [right=of z,label=below:w] { };
  \path (alpha) edge [connect] (theta)
        (theta) edge [connect] (z)
        (z) edge [connect] (w)
        (beta) edge [connect] (w);
  \node[rectangle, inner sep=0mm, fit= (z) (w),label=below right:N, xshift=13mm] {};
  \node[rectangle, inner sep=4.4mm,draw=black!100, fit= (z) (w)] {};
  \node[rectangle, inner sep=4.6mm, fit= (z) (w),label=below right:M, xshift=12.5mm] {};
  \node[rectangle, inner sep=9mm, draw=black!100, fit = (theta) (z) (w)] {};
\end{tikzpicture}
\end{figure}

\end{frame}

\begin{frame}{Latent Dirichlet Allocation: Some Details}

\begin{itemize}
\item We use $K = 30$ topics for our resulting model.
\item What we ultimately care about are $\theta$ and $\beta$.
\item $\theta$ corresponds to how likely a topic is to appear in a document.
\item $\beta$ corresponds to how likely each word is to be associated with each topic.
\end{itemize}

\end{frame}

\begin{frame}{Latent Dirichlet Allocation: Some Details}

\begin{itemize}
\item Some questions we want to answer with LDA
\begin{itemize}
\item Can we see sensible topics in the LDA output?
\item Who is associated most with each issue?
\item Do the topics proportions for each email line up with real world events?
\item Can we predict the source of a given email?
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{Some Interesting Topics}

\begin{table}[h]
\centering
\label{topic_top_words}
\scalebox{0.8}{
\begin{tabular}{cccccc}
\hline
Israel & Elections & Libya & Afghanistan & Int. Dev. & Obama \\
\hline
israel & democrat & libya & afghanistan & develop & presid \\
isra & republican & secur & pakistan & state & obama \\
peace & american & travel & afghan & support & said \\
palestinian & polit & libyan & militari & global & hous \\
netanyahu & parti & iraq & general & program & white \\
east & elect & embassi & karzai & effort & administr \\
negoti & obama & attack & offici & intern & polici \\
arab & percent & kill & war & work & aid \\
state & candid & march & forc & includ & advis

\end{tabular}
}
\end{table}

\end{frame}

\begin{frame}{Topic Importance Over Time}

\begin{figure}[h]
\centering
\subfigure[Israel]{\label{fig:t1}\includegraphics[width=29mm]{../images/time_plot6.pdf}}
\subfigure[Elections]{\label{fig:t2}\includegraphics[width=29mm]{../images/time_plot11.pdf}}
\subfigure[Libya]{\label{fig:t3}\includegraphics[width=29mm]{../images/time_plot18.pdf}}
\subfigure[Afghanistan]{\label{fig:t4}\includegraphics[width=29mm]{../images/time_plot27.pdf}}
\subfigure[Int. Dev.]{\label{fig:t5}\includegraphics[width=29mm]{../images/time_plot23.pdf}}
\subfigure[Obama]{\label{fig:t6}\includegraphics[width=29mm]{../images/time_plot25.pdf}}
\label{fig:topic_time_plots}
\end{figure}

\end{frame}

\begin{frame}{Who Says What?}

\begin{table}[h]
\centering
\label{topic_sources}
\scalebox{0.5}{
\begin{tabular}{cccccc}
\hline
Israel & Elections & Libya & Afghanistan & Int. Dev. & Obama \\
\hline
Sidney Blumenthal & SB & Huma Abedin & Judith McHale & JM & SB \\
(SB) & & (HA) & (JM) & & \\
\noalign{\vskip 7mm}
HA & Philippe Reines & Wendy Sherman & HA & Melanne Verveer & PR \\
& (PR) & (WS) & & (MV) & \\
\noalign{\vskip 7mm}
Jake Sullivan & JM & JS & MV & Anne-Marie & Cherly Mills \\
(JS) & & & & Slaughter (AMS) & (CM)\\
\noalign{\vskip 7mm}
AMS & CM & Monica Hanley & JS & CM & HA \\
& & (MH) & & & \\
\noalign{\vskip 7mm}
Hillary Clinton & HA & SB & Richard Verma & JS & HC \\
(HC) & & & (RV) & & \\

\end{tabular}
}
\end{table}

\end{frame}

\begin{frame}{Multinomial Logistic Regression}

\begin{itemize}
\item We next want to ask can we predict who said what?
\item To answer this question we employ multinomial regression.
\item We take a subset of people who sent emails, only those who have more that 100 emails sent and try prediction for these.
\end{itemize}

\end{frame}

\begin{frame}{Multinomial Logistic Regression Results}

\begin{table}[h]
\centering
\label{multinomial_results}
\scalebox{0.7}{
\begin{tabular}{ccc}
\hline
Source & Success Rate & Testing Observations \\
Hillary Clinton & 0.74 & 849 \\
Philippe Reines & 0.06 & 34 \\
Claire Coleman & 0.52 & 27 \\
Lauren Jiloty & 0.42 & 71 \\
Huma Abedin & 0.36 & 376 \\
Jake Sullivan & 0.23 & 410 \\
Sidney Blumenthal & 0.29 & 87 \\
Cherly Mills & 0.31 & 491 \\
Anne-Marie Slaughter & 0.36 & 42 \\
Monica Hanley & 0.09 & 53 \\
Judith McHale & 0.25 & 28 \\
Robert Russo & 0.00 & 10 \\
Richard Verma & 0.26 & 19 \\
Wendy Sherman & 0.08 & 12 \\
Melanne Verveer & 0.30 & 30 \\
Lona Valmoro & 0.34 & 41
\end{tabular}
}
\end{table}

\end{frame}

>>>>

% \documentclass[13pt]{beamer}
\documentclass[handout,13pt]{beamer}

@@

\usetheme{Madrid}
\usecolortheme{beaver}

@@

\begin{frame}{Data}
\begin{itemize}
\item From 2008 to 2013, Hillary Clinton and some of her staff used a private email server for much of their communication
\item In response to FOIA requests, the State Department has released pdfs of roughly 30,000 of these emails, with roughly 8,000 sent by Clinton
\item We have a dataset of all these released emails, including sender and receiver information\footnote{Thanks to Ben Hamner and the WSJ for making their code and data available, respectively.}
\pause\item Bag-of-words model
  \begin{itemize}
  \item Stemming: combining words with the same ``root''
  \item Remove extremely rare words and extremely common (stop) words
  \end{itemize}
\item After cleaning, roughly 28,000 emails and a total vocabulary of over 3,000 words
\end{itemize}
\end{frame}

\begin{frame}{Visualization}
\begin{itemize}
\item In order to visualize these emails, we first need a measure of distance
\item Cosine similarity:
  \begin{equation*}
  \cos(\phi) = \frac{<V, W>}{\norm{V}\norm{W}}
  \end{equation*}
\item We can then use classical multi-dimensional scaling to visualize this data in lower dimensions
\end{itemize}
\end{frame}

\begin{frame}{Visualization}
\begin{figure}[h]
\centering
\subfigure[2D]{\label{fig:mds_2d}\includegraphics[width=0.44\linewidth]{../images/mds_2d.pdf}}
\subfigure[3D]{\label{fig:mds_3d}\includegraphics[width=0.54\linewidth]{../images/mds_3d_1.pdf}}
\label{fig:mds}
\end{figure}
\end{frame}

\begin{frame}{Spectral Clustering}
\begin{itemize}
\item Cluster points using their pairwise similarity
\item Computer the first $K$ eigenvectors of the normalized Laplacian
  \begin{equation*}
  L = I - D^{-1/2} S D^{-1/2}
  \end{equation*}
\item Perform $k$-means clustering on the resulting eigenvectors
\end{itemize}
\end{frame}

\begin{frame}{Spectral Clustering}
\begin{figure}[h]
\centering
\subfigure[2D]{\label{fig:cluster_2d}\includegraphics[width=0.44\linewidth]{../images/mds_2d_cluster.pdf}}
\subfigure[3D]{\label{fig:cluster_3d}\includegraphics[width=0.54\linewidth]{../images/mds_3d_cluster_1.pdf}}
\label{fig:cluster}
\end{figure}
\end{frame}

\begin{frame}{Latent Dirichlet Allocation}
\begin{figure}
\centering
\begin{tikzpicture}[scale=1.0, every node/.style={transform shape}]
\tikzstyle{main}=[circle, minimum size = 10mm, thick, draw =black!80, node distance = 16mm]
\tikzstyle{connect}=[-latex, thick]
\tikzstyle{box}=[rectangle, draw=black!100]
  \node[main] (alpha) {$\alpha$};
  \node[main] (theta) [right=of alpha] {$\theta$};
  \node[main] (z) [right=of theta] {$z$};
  \node[main] (beta) [above=of z] {$\beta$};
  \node[main, fill=black!10] (w) [right=of z] {$w$};
  \edge {alpha} {theta};
  \edge {theta} {z};
  \edge {z,beta} {w};
  \plate {words} {(z)(w)} {$N_m$};
  \plate {docs} {(theta)(z)(w)(words.south east)(words.north west)} {$M$};
\end{tikzpicture}
\end{figure}
\end{frame}

\begin{frame}{Latent Dirichlet Allocation: Some Details}
\begin{itemize}
\item We use $K = 30$ topics for our resulting model.
\item What we ultimately care about are $\theta$ and $\beta$.
\item $\theta$ corresponds to how likely a topic is to appear in a document.
\item $\beta$ corresponds to how likely each word is to be associated with each topic.
\end{itemize}
\end{frame}

\begin{frame}{Latent Dirichlet Allocation: Some Details}
\begin{itemize}
\item Some questions we want to answer with LDA
\begin{itemize}
\item Can we see sensible topics in the LDA output?
\item Who is associated most with each issue?
\item Do the topics proportions for each email line up with real world events?
\item Can we predict the source of a given email?
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Some Interesting Topics}
\begin{table}[h]
\centering
\label{topic_top_words}
\scalebox{0.8}{
\begin{tabular}{cccccc}
\toprule
Israel & Elections & Libya & Afghanistan & Int. Dev. & Obama \\
\midrule
israel & democrat & libya & afghanistan & develop & presid \\
isra & republican & secur & pakistan & state & obama \\
peace & american & travel & afghan & support & said \\
palestinian & polit & libyan & militari & global & hous \\
netanyahu & parti & iraq & general & program & white \\
east & elect & embassi & karzai & effort & administr \\
negoti & obama & attack & offici & intern & polici \\
arab & percent & kill & war & work & aid \\
state & candid & march & forc & includ & advis \\
\bottomrule
\end{tabular}
}
\end{table}
\end{frame}

\begin{frame}{Topic Importance Over Time}
\begin{figure}[h]
\centering
\subfigure[Israel]{\label{fig:t1}\includegraphics[width=0.25\linewidth]{../images/time_plot6.pdf}}
\subfigure[Elections]{\label{fig:t2}\includegraphics[width=0.25\linewidth]{../images/time_plot11.pdf}}
\subfigure[Libya]{\label{fig:t3}\includegraphics[width=0.25\linewidth]{../images/time_plot18.pdf}}
\subfigure[Afghanistan]{\label{fig:t4}\includegraphics[width=0.25\linewidth]{../images/time_plot27.pdf}}
\subfigure[Int. Dev.]{\label{fig:t5}\includegraphics[width=0.25\linewidth]{../images/time_plot23.pdf}}
\subfigure[Obama]{\label{fig:t6}\includegraphics[width=0.25\linewidth]{../images/time_plot25.pdf}}
\label{fig:topic_time_plots}
\end{figure}
\end{frame}

\begin{frame}{Who Says What?}
\begin{table}[h]
\centering
\label{topic_sources}
\scalebox{0.58}{
\begin{tabular}{cccccc}
\toprule
Israel & Elections & Libya & Afghanistan & Int. Dev. & Obama \\
\midrule
Sidney Blumenthal & SB & Huma Abedin & Judith McHale & JM & SB \\
(SB) & & (HA) & (JM) & & \\
\noalign{\vskip 7mm}
HA & Philippe Reines & Wendy Sherman & HA & Melanne Verveer & PR \\
& (PR) & (WS) & & (MV) & \\
\noalign{\vskip 7mm}
Jake Sullivan & JM & JS & MV & Anne-Marie & Cherly Mills \\
(JS) & & & & Slaughter (AMS) & (CM)\\
\noalign{\vskip 7mm}
AMS & CM & Monica Hanley & JS & CM & HA \\
& & (MH) & & & \\
\noalign{\vskip 7mm}
Hillary Clinton & HA & SB & Richard Verma & JS & HC \\
(HC) & & & (RV) & & \\
\bottomrule
\end{tabular}
}
\end{table}
\end{frame}

\begin{frame}{Multinomial Logistic Regression}
\begin{itemize}
\item We next want to ask can we predict who said what?
\item To answer this question we employ multinomial regression.
\item We take a subset of people who sent emails, only those who have more that 100 emails sent and try prediction for these.
\end{itemize}
\end{frame}

\begin{frame}{Multinomial Logistic Regression Results}
\begin{table}[h]
\centering
\label{multinomial_results}
\scalebox{0.7}{
\begin{tabular}{ccc}
\toprule
Source & Success Rate & Testing Observations \\
\midrule
Hillary Clinton & 0.74 & 849 \\
Philippe Reines & 0.06 & 34 \\
Claire Coleman & 0.52 & 27 \\
Lauren Jiloty & 0.42 & 71 \\
Huma Abedin & 0.36 & 376 \\
Jake Sullivan & 0.23 & 410 \\
Sidney Blumenthal & 0.29 & 87 \\
Cherly Mills & 0.31 & 491 \\
Anne-Marie Slaughter & 0.36 & 42 \\
Monica Hanley & 0.09 & 53 \\
Judith McHale & 0.25 & 28 \\
Robert Russo & 0.00 & 10 \\
Richard Verma & 0.26 & 19 \\
Wendy Sherman & 0.08 & 12 \\
Melanne Verveer & 0.30 & 30 \\
Lona Valmoro & 0.34 & 41 \\
\bottomrule
\end{tabular}
}
\end{table}
\end{frame}
